---
output: github_document
---

# panelcert

panelcert screens fixed-effect panel designs for concentrated identifying variation, leverage, measurement error, and TWFE heterogeneity, and supplies exact contrast inference where Gaussian approximations are fragile.

[![R-CMD-check](https://github.com/profsms/panelcert/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/profsms/panelcert/actions/workflows/R-CMD-check.yaml)
[![Cross-language parity](https://github.com/profsms/panelcert/actions/workflows/parity.yaml/badge.svg)](https://github.com/profsms/panelcert/actions/workflows/parity.yaml)
[![Codecov](https://codecov.io/gh/profsms/panelcert/branch/main/graph/badge.svg)](https://app.codecov.io/gh/profsms/panelcert)

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Install

Install the current release from GitHub:

```r
install.packages("remotes")
remotes::install_github("profsms/panelcert@v0.8.0")
```

The r-universe build will be installable with `install.packages("panelcert", repos = c("https://profsms.r-universe.dev", "https://cloud.r-project.org"))`. CRAN installation will be documented after acceptance.

## Quickstart

```{r quickstart}
library(panelcert); data(grunfeld)
applicable(grunfeld$capital, grunfeld$firm, grunfeld$year, controls = grunfeld$value)
report <- cycle_report(grunfeld$invest, grunfeld$capital, grunfeld$firm, grunfeld$year, controls = grunfeld$value, interval = FALSE)
c(report$verdict, lambda_score = report$statistic$lambda_score, kappa = report$statistic$kappa)
```

## Entry Points

- `leverage_report(y, x, unit, time)` is the diffuse-regime diagnostic. It checks treatment concentration and leverage balance and compares df-corrected and HC0-HC3 inference.
- `cycle_report(y, x, unit, time)` is the concentrated-regime workflow from Paper A. It reports capture, granularity, an exact sign-flip test, and an exact confidence set.
- `eiv_adequacy(y, x, unit, time, ...)` screens continuous-regressor measurement error under the article's exact-normal mapping and certificate. Supply `reliability_lower` and `gamma_lambda` when the reliability pilot is noisy; otherwise certification is explicitly conditional on treating the computed reliability as known/consistent.
- `breakdown_reliability(...)` returns the descriptive point threshold, while `certified_breakdown_reliability(...)` adds the coefficient-uncertainty budget in closed form.
- `reliability_from_repeats(first, second)` estimates within reliability from identically projected repeated measurements; use `method = "equal_variance"` only when that added restriction is defensible.
- `twfe_adequacy(y, unit, time, first_treat)` reports the saturated group-time
  envelope, the cohort/event/additive restricted ladder, and a signed
  directional plug-in. Its formal decision uses a multiplier lower test and a
  covariance-aware projected-Wald upper certificate for a prespecified class.
  Direct CR1 cluster-score normalization is the default; `cluster = "ar1"`,
  `cluster = "iid"`, and a positive user-supplied `psi` are sensitivity routes.

Default `cycle_report()` printing is concise. Detailed caveats remain available
in `report$notes`, `show_notes(report)`, or `print(report, notes = TRUE)`.

`adequacy_row(x, unit, time, y = NULL)` composes the design and capture diagnostics into one flat record for prevalence screens.

The regression diagnostics accept `controls =` for numeric nuisance covariates. Exact cycle packing supports one linearly independent control and constructs locally annihilating 2-by-3 supports; it never substitutes globally residualized outcomes for valid disjoint contrasts.

`applicable(x, unit, time)` is the outcome-free pre-flight check; it reports structural binary-treatment granularity failures before packing.

## Packing Methods

- `"structured"` exploits four-cycle structure in dense panels and 2-by-3 local projections when one continuous control is supplied.
- `"sparse"` is the scalable method for AKM-style mobility networks.
- `"greedy"` is a cheap lower bound that keeps the best of four deterministic DFS traversals.

## Verdicts

- `POINT_PASS` is a descriptive pass at the requested finite-design threshold; it is not a theorem-level certificate.
- `FLAGGED` marks a concentration or adequacy failure for which conventional inference is not licensed.
- `INCONCLUSIVE` is returned when the enumeration floor `2^(1-effective_C)` exceeds `alpha`. Its `reason` distinguishes the coding-invariant binary-treatment floor based on `min(n_treated, n_untreated)`, which repacking cannot fix, from a selected packing that may be improved.

Measurement-error reports can additionally return `CERTIFIED` when the supplied reliability lower bound clears the certified breakdown. The report prints `gamma_beta`, `gamma_lambda`, and their total false-certification bound; the legacy keyword `gamma` denotes `gamma_beta`.

TWFE reports use `CERTIFIED` only when the covariance-aware upper bound for the
selected class lies below the exact-normal threshold, `FLAGGED` when the
regular one-sided lower bound lies above it, and `INCONCLUSIVE` otherwise. A
rank-deficient projected covariance cannot support an upper certificate. Set
`heterogeneity_class` before examining the outcome; `q_band` requests the more
general but less powerful confidence-ball fallback without changing the main
verdict.

## Design-Only Use

`design_summary`, `cycle_capture`, and `adequacy_row(..., y = NULL)` require only `(x, D, controls)`, not an outcome. In particular, `rho`, `V_n`, `lambda_n`, `n_eff`, and `kappa_C` are available before estimation.

## Julia Package And Data

The sibling Julia package is [PanelAdequacy.jl](https://github.com/profsms/PanelAdequacy.jl). Fixed CSV designs and CI enforce agreement to `1e-12` on numerical outputs and exact agreement on integer fields, verdicts, and packed supports.

Dataset access follows language conventions: R exposes bundled panels through `LazyData`; Julia exposes `datasets()`, `datapath()`, and `load_dataset()`. Redistributable paper panels, including V-Dem, repeated-report twins, the 2,284-county minimum-wage application, the 2,840-municipality Brazil property-tax reanalysis, and the public-domain canonical 11-firm Grunfeld showcase, are bundled for offline replication. `inst/DATA_SOURCES.md` records pinned provenance and explains why the KSS test extract uses a checksum-pinned direct download instead.

## Citation

Please cite the corresponding working paper when a diagnostic is used:

- Halkiewicz, Stanislaw M. S. (2026). *Exact Inference in Fixed-Effect Regressions with Concentrated Identifying Variation*.
- Halkiewicz, Stanislaw M. S. (2026). *Breakdown Reliability for Saturated Fixed-Effect Inference*.
- Halkiewicz, Stanislaw M. S. (2026). *Positive Weights Do Not Certify TWFE Inference*.

Run `citation("panelcert")` for machine-readable package metadata.
