---
title: "Paper A: exact contrast inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Paper A: exact contrast inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(7)
```

## Design, contrasts, and exactness

Paper A separates two questions. Treatment concentration (`lambda_n`) warns
when a Gaussian approximation may be unreliable. Nuisance-annihilating
contrasts then provide finite-sample sign-flip inference under the paper's
joint sign-invariance assumption. Exactness comes from that assumption and
verified nuisance annihilation, not from a small value of `lambda_n`.

```{r}
library(panelcert)
N <- 24
worker <- rep(seq_len(N), each = 2)
year <- rep(1999:2000, times = N)
x <- rnorm(2 * N)
y <- 0.5 * x + rnorm(2 * N)

out <- cycle_report(y, x, worker, year, nflips = 999,
                    interval = FALSE)
out
```

`cycle_contrasts()` constructs two-way cycle systems. For worker--firm--year
and other multiway designs, construct design-only supports explicitly and let
`contrast_system()` verify every fixed-effect dimension:

```{r}
worker <- c(1, 1, 2, 2)
firm <- c(10, 10, 20, 20)
year <- c(1999, 2001, 1999, 2001)
x <- c(0, 0, 0, 1)

cs <- contrast_system(
  x,
  fe_levels = list(worker, firm, year),
  rows = list(1:4),
  weights = list(c(-1, 1, 1, -1)),
  blocks = worker
)
support_compatibility(cs, worker)
```

With clustered errors, pass dependence-block ids to construction and
inference. Every treatment-loaded support must contain complete blocks, and a
block cannot be split between supports. `cycle_report(..., blocks = worker)`
also reports treatment and realized-score concentration at the worker level.

## Interpretation

`kappa` is always the observable fraction of within-treatment variation
captured by the contrast system. It is a Pitman efficiency only when both power
conditions in Paper A hold. `POINT_PASS` means only that the finite design lies
below a user-selected concentration warning threshold; it is not a theorem-
level certificate. A `FLAGGED` design is precisely where the exact contrast
interval is most relevant.
