---
title: "Concentrated design: diagnostics to exact inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Concentrated design: diagnostics to exact inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(7)
```

## Bundled application

Paper A's return-panel application is bundled as `fscore`. The specification
uses firm and country-year fixed effects, so every number below is reproducible
without a download.

```{r}
library(panelcert)
data(fscore)
cy <- interaction(fscore$country, fscore$year, drop = TRUE)
y <- log1p(fscore$ret)
x <- fscore$fscore
```

## Diagnose concentration

The design summary requires no outcome. `lambda_n` is the largest observation's
share of within-treatment variation and `n_eff` is its inverse-Herfindahl
effective support.

```{r}
d <- design_summary(fscore$uid, cy, x = x)
d
```

The one-call screen adds the greedy lower bound and the designed packing. It
can be used for a prevalence exercise before outcomes are collected.

```{r}
screen <- adequacy_row(x, fscore$uid, cy, method = "structured")
screen[c("Vn", "lambda_n", "n_eff", "kappa_greedy",
         "kappa_designed", "se_price")]
```

## Capture and granularity

The structured packing uses disjoint four-cycles. `kappa` is the observable
share of within-treatment variation captured by its annihilating contrasts;
it is a Pitman efficiency only under Paper A's two additional power conditions.

```{r}
cs <- cycle_contrasts(x, fscore$uid, cy, method = "structured")
c(C = length(cs$rows), kappa = cs$kappa,
  effective_C = length(panelcert:::.active_supports(cs)))
```

Every contrast satisfies `v'D = 0`. Consequently `v'x_tilde = v'x` exactly:
packing is ranked on raw `x`, while iterative residualization is used only for
the denominator `V_n`. Both package implementations regression-test this
identity and the selected supports.

## Exact interval

The report adds realized score concentration and the pass/flag/inconclusive
verdict. The interval is then obtained by sign-flip inversion. The reduced
Monte Carlo and grid sizes below keep vignette builds fast; replication runs
should increase them.

```{r}
report <- cycle_report(y, x, fscore$uid, cy, method = "structured",
                       nflips = 1999, interval = FALSE)
report

ci <- signflip_interval(y, cs, alpha = 0.05, nflips = 1999,
                        ngrid = 2401)
unlist(ci[c("lo", "hi", "beta_tilde", "effective_C", "kappa")])
```

Exactness requires joint sign invariance at the dependence-block level and
supports that contain whole blocks. For clustered errors, pass `blocks` to
`cycle_contrasts()`, `cycle_report()`, and `signflip_interval()`; the package
rejects an incompatible packing instead of silently asserting exactness.
