---
title: "Module A: variance estimators under fixed-effect saturation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Module A: variance estimators under fixed-effect saturation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed (see panelcert.Rmd).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(2)
```

## The problem

As fixed effects saturate (`rho = d_K/n` grows), conventional variance
estimators split in opposite directions (Paper A, Theorem 3.3):

- **HC0** is biased downward by the factor `(1 - rho)`: t-tests over-reject,
  with asymptotic size `2(1 - pnorm(z * sqrt(1 - rho)))` — about 9% at
  `rho = 0.25` and 17% at `rho = 0.5` for a nominal 5% test;
- **HC3** over-corrects by exactly the inverse factor `1/(1 - rho)`: size
  drops below 1% at `rho = 0.5`, and confidence intervals are artificially
  wide;
- **HC2 / leave-one-out** (and HC1 under uniform leverage) is the estimator
  that stays exact.

`leverage_report()` reproduces your FE regression via Frisch–Waugh, computes
the full hierarchy of standard errors with the exact full-regression leverage
`H_ii = (P_K)_ii + Xt_i^2 / (X'MX)`, and reports whether the estimator choice
materially changes inference on your design.

## A saturated design

```{r}
library(panelcert)
N <- 12; T <- 5
unit <- rep(1:N, each = T); time <- rep(1:T, times = N)
n <- N * T
x <- rnorm(n) + 0.3 * unit
y <- 0.8 * x + rnorm(n)
leverage_report(y, x, unit, time)
```

At `rho = (N + T - 1)/n` near 0.27, the HC0 and HC3 standard errors bracket
the leave-one-out one by visibly asymmetric factors, and the report's implied
size quantifies what an HC0 user would actually suffer. The `breakdown
threshold` line is the saturation `rho_dagger` (about 0.296 at
`alpha = delta = 0.05`) at which HC0-based inference exits the tolerance.

## Reading the flags

Two notes deserve attention when they appear:

- the **HC3 over-correction flag** fires whenever `rho > 0.1` (Paper A's rule
  of thumb): switching to leave-one-out recovers the honestly narrower
  interval;
- the **leverage non-uniformity flag** (`hmax/hmin` large) marks designs —
  small country-year cells, "anchor" units — where HC1 silently diverges from
  leave-one-out and only the latter remains reliable (Paper A, Remark 3.4).

Fitted-model ingestion works the same way:

```{r, eval = requireNamespace("fixest", quietly = TRUE)}
d <- data.frame(y = y, x = x, unit = unit, time = time)
m <- fixest::feols(y ~ x | unit + time, data = d)
leverage_report(m)$verdict
```

## The empirical application (Paper A), reproduced from bundled data

Paper A's firm panel ships with the package as `fscore`: Piotroski F-Scores and
one-year-ahead returns for Warsaw, Budapest and Prague listings, 2010--2024
(217 firm-years, 19 firms). The preferred specification --- log return on the
F-Score, firm and year fixed effects --- runs directly off it:

```{r}
library(panelcert)
r <- leverage_report(log1p(fscore$ret), fscore$fscore, fscore$uid, fscore$year)
c(beta = round(r$statistic$beta, 5),
  HC0_LO = round(r$statistic$se_hc0 / r$statistic$se_hc2, 3),
  HC3_LO = round(r$statistic$se_hc3 / r$statistic$se_hc2, 3),
  rho = round(r$design$rho, 3))
```

The empirical HC0/LO and HC3/LO ratios track the theoretical `sqrt(1 - rho)`
and `1/sqrt(1 - rho)` almost exactly, and the F-Score coefficient is
indistinguishable from zero. The most favourable specification is Poland alone,
whose `|t|` is still only 0.61:

```{r}
pl <- fscore$country == "Poland"
e <- leverage_report(log1p(fscore$ret[pl]), fscore$fscore[pl],
                     fscore$uid[pl], fscore$year[pl])
round(abs(e$statistic$beta / e$statistic$se_hc2), 2)
```

The country-year-fixed-effect specification is the leverage cautionary tale of
the flag above: each firm sits in exactly one country, so the firm +
country-year design splits into three disconnected country blocks and the
realized FE dimension is `d_K = 61`, which `leverage_report()` computes from the
actual design via union-find rather than a nominal `19 + 45 - 1` count. Pass
`paste(fscore$country, fscore$year)` as the time id to reproduce it.
`data(package = "panelcert")` lists the bundled datasets.
