---
title: "Module B: measurement-error adequacy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Module B: measurement-error adequacy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed (see panelcert.Rmd).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(3)
```

## The problem

Fixed-effect saturation removes signal from the regressor while (largely)
preserving measurement noise — the Griliches–Hausman amplification. Paper B
shows the resulting size distortion of the naive t-test is governed by one
number, the **within reliability**

`lambda_hat = 1 - mean(sigma_nu^2) * (n - d_K) / tau_star2`,

the share of post-projection variation in the observed regressor that is
signal. The feasible non-centrality is
`|eta| = (|beta0|/sigma) * (1 - lambda) * sqrt(tau_star2)`, and naive
inference is `delta`-adequate iff `|eta| <= eta_dagger(alpha, delta)` — the
**exact-inversion** threshold (about 0.652 at `alpha = delta = 0.05`). The
quadratic closed form `sqrt(delta/(z*phi(z)))` is its accurate companion; a
linear expansion is spurious (the size is even in `eta`) and must not be used.

## Noise inputs

Three pathways, in decreasing order of directness:

1. **published measurement-model posteriors** (V-Dem-style):
   `codelow`/`codehigh` interval bounds, converted by
   `reliability_from_interval()`;
2. **validation-study reliability ratios** (PSID-style):
   `reliability_from_ratio(r, within_sd)`;
3. **direct** `sigma_nu` or `reliability` values, including a sensitivity grid.

## The honesty machinery, demonstrated

The most natural implementation — plugging the attenuated `beta*` from the
regression table into the threshold — is biased toward *certifying failing
specifications*, by exactly the factor `lambda` whose smallness is the reason
to run the diagnostic (Paper B, Proposition on pilots). The package defaults
to the corrected pilot at its upper confidence bound; the naive form is
opt-in and labelled. On a design built to fail:

```{r}
library(panelcert)
N <- 100; T <- 8
unit <- rep(1:N, each = T); time <- rep(1:T, times = N)
n <- N * T
x_true <- 0.5 * rnorm(n)
x_obs <- x_true + rnorm(n, sd = 0.45)   # within reliability ~ 0.55
y <- 1.0 * x_true + rnorm(n)

r_default <- eiv_adequacy(y, x_obs, unit, time, sigma_nu = 0.45)
r_naive <- eiv_adequacy(y, x_obs, unit, time, sigma_nu = 0.45, pilot = "naive")
c(default = r_default$verdict, naive = r_naive$verdict)
```

The naive pilot understates `|eta|` by the factor `lambda` and can certify a
specification whose true size is several times nominal — in Paper B's V-Dem
application it certifies the legislative-constraints index whose exact size is
14%. The corrected default flags it.

```{r}
r_default
```

Alongside the verdict, the report carries the **breakdown reliability**
`lambda_dagger` — how reliable the regressor would have to be for naive
inference to pass — and the `sqrt(lambda)` **power tax**, which applies even
to certified specifications.

## The empirical poles (Paper B), reproduced from bundled data

The V-Dem democracy–growth panel ships with the package as `vdem` (the
measurement-error SDs are published data). Paper B's two-pole result runs
directly off it — no download:

```{r}
# Aggregate polyarchy: certified (lambda ~ 0.90, implied size ~5.6%)
p <- vdem[complete.cases(vdem$ly, vdem$v2x_polyarchy, vdem$v2x_polyarchy_sd), ]
eiv_adequacy(p$ly, p$v2x_polyarchy, p$iso, p$year, sigma_nu = p$v2x_polyarchy_sd)

# Legislative constraints: flagged (lambda ~ 0.55, implied size ~14%)
l <- vdem[complete.cases(vdem$ly, vdem$v2xlg_legcon, vdem$v2xlg_legcon_sd), ]
eiv_adequacy(l$ly, l$v2xlg_legcon, l$iso, l$year,
             sigma_nu = l$v2xlg_legcon_sd)$verdict
```

The second application ships as `psid` (Cornwell–Rupert earnings): the
within transformation alone moves the log wage from certified at the level
reliability 0.82 to flagged across the validated within band 0.50–0.65.

```{r}
eiv_adequacy(psid$wks, psid$lwage, psid$id, psid$year,
             reliability = 0.65, pilot = "point")$verdict
```

Every number in Paper B's tables is reproduced this way; `data(package =
"panelcert")` lists the bundled datasets.
