---
title: "panelcert: certifying naive panel inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{panelcert: certifying naive panel inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed. Before the
# package is built (e.g. knitting straight from the source tree) chunks are
# shown but not run, so the vignette never errors on a missing `panelcert`.
# During R CMD check the package IS available, so everything runs as normal.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(1)
```

## What the package does

`panelcert` diagnoses whether conventional fixed-effect inference is reliable
and implements Paper A's exact nuisance-annihilating contrast alternative.
The diagnostic modules consume a fitted model (`fixest::feols`, `plm::plm`,
`lm`) or raw panel; the contrast engine works from the design and outcome.
Reports use four plain-language verdicts:

- **CERTIFIED at tolerance delta** — the worst-case size of the nominal-alpha
  test stays below `alpha + delta`;
- **FLAGGED (implied size X%)** — it does not, and the report quantifies the
  failure and points to the appropriate remedy;
- **POINT PASS** — a plug-in or finite-threshold comparison passes, but is not
  a formal confidence-band certificate;
- **INCONCLUSIVE** — a pre-outcome design statistic that needs a pilot input.

The current source map is:

| Source | Problem | Headline output |
|---|---|---|
| Paper A | concentrated identifying variation | `lambda_n`, `kappa`, exact sign-flip interval |
| Paper B | classical measurement error | within reliability `lambda_hat` |
| Paper C | TWFE heterogeneity in staggered DiD | design statistic `Gamma` |
| Diffuse companion | variance-estimator choice under FE saturation | HC0–HC3 hierarchy, leverage |

## Bundled data — the paper applications run offline

The small reference panels ship with the package. After `library(panelcert)`
the datasets are available by name:

```{r}
library(panelcert)
data(package = "panelcert")$results[, "Item"]   # includes brazil, castle, divorce, minimum_wage
twfe_adequacy(castle$y, castle$uid, castle$tid, castle$ft,
              heterogeneity_class = "additive")$verdict
```

Paper A's F-score workflow, Papers B and C, and the diffuse companion use these
objects. The public KSS worker--firm testing extract is intentionally not
vendored; Paper A's replication code downloads it from the source repository.

## The shared design summary

Every module rests on the same design primitives: the fixed-effect dimension
`d_K` (union-find on the unit–time graph), the saturation `rho = d_K/n`, and
the two-way within transformation.

```{r}
library(panelcert)
unit <- rep(1:40, each = 12)
time <- rep(1:12, times = 40)
design_summary(unit, time)
```

## A sixty-second tour

Paper A exact inference in a balanced worker-time panel:

```{r}
set.seed(3)
n <- length(unit)
x <- rnorm(n)
y <- 0.4 * x + rnorm(n)
cycle_report(y, x, unit, time, nflips = 199, interval = FALSE)
```

Module B on a panel with a noisy regressor (measurement-error SD known, e.g.
from a validation study or a published measurement model):

```{r}
x_true <- rnorm(n) + 0.2 * unit          # true regressor
x_obs <- x_true + rnorm(n, sd = 0.3)     # observed with error
y <- 0.6 * x_true + rnorm(n)
eiv_adequacy(y, x_obs, unit, time, sigma_nu = 0.3)
```

Module C, before any outcome exists — vetting a staggered adoption design:

```{r}
ft <- ifelse(unit <= 15, 4, ifelse(unit <= 30, 8, NA))
twfe_design(unit, time, ft)
```

The verdict logic is *correct by default*: Module B evaluates the certificate
at the upper confidence bound of the attenuation-corrected pilot (the naive
attenuated pilot is available only as an explicitly labelled opt-in); Module C
uses the shrinkage-corrected cohort dispersion and computes the cluster
variance-inflation factor from the realized design rather than assuming its
direction. These defaults guard against specific errors the source papers
document; the module vignettes explain each one.

## Where the numbers come from

All thresholds derive from one limit experiment: the studentized statistic is
asymptotically `N(eta, 1)`, the two-sided size is even in `eta` with leading
distortion `z * phi(z) * eta^2` (quadratic — a linear expansion is spurious and
badly anti-conservative), and the operational threshold `eta_dagger(alpha,
delta)` is the exact-inversion root of `size(eta) = alpha + delta`, about 0.652
at `alpha = delta = 0.05`. The empirical applications in the source papers
(V-Dem democracy–growth, castle-doctrine, minimum-wage and no-fault-divorce
DiD) are reproduced to reference tolerance by this package's test suite, in
lockstep with the Julia reference implementation.
