---
title: "TWFE-heterogeneity adequacy for staggered DiD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TWFE-heterogeneity adequacy for staggered DiD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed (see panelcert.Rmd).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(4)
```

## The design statistic, before any outcome exists

Under heterogeneous treatment effects, the TWFE coefficient is a weighted
average with weights `w` that need not be convex. The accompanying article shows the *size
distortion* of the TWFE t-test is governed by

`Gamma = sqrt(N1) * ||w - u||`,

the scaled distance of the weights from uniform — a **pure design statistic**,
computable from the adoption pattern alone. `Gamma = 0` exactly for clean
block designs; it grows with staggering and negative weights. Crucially,
`Gamma` is invariant to replicating the panel while the absolute weighting
bias shrinks — so a design can look safe to estimand-level diagnostics while
its t-test stays arbitrarily distorted.

```{r}
library(panelcert)
N <- 48; T <- 12
unit <- rep(1:N, each = T); time <- rep(1:T, times = N)

# a block design: one adoption date, never-treated reservoir
ft_block <- ifelse(unit <= 16, 5, NA)
twfe_design(unit, time, ft_block)$statistic$Gamma

# three cohorts, no never-treated reservoir
ft_stag <- ifelse(unit <= 16, 3, ifelse(unit <= 32, 7, 11))
twfe_design(unit, time, ft_stag)
```

Run this before committing to a design: the report's breakdown value is the
largest heterogeneity-to-noise ratio `c/sigma` the design tolerates.

## The inference layer

With an outcome, `twfe_adequacy()` estimates heterogeneity in a nested ladder:
cohort, event time, additive cohort plus event time, and the saturated
post-treatment group-time class. In a balanced panel the last class spans every
treated-cell profile relevant to TWFE weighting, so `Gamma_gt = Gamma`. The
function maps the estimated profile to two distinct quantities:

- the **worst-case size envelope** maximizes the mean shift over that calibrated
  class. It says what the design can certify uniformly; it is an upper bound,
  not a point prediction of the test's realized rejection probability;
- the **directional plug-in** inserts the estimated signed group-time profile.
  It is orientation-sensitive corroboration, not a uniform certificate.

The default scale is direct CR1. The function forms the unit cluster scores,
sets `psi_direct` to their CR1 variance relative to the homoskedastic scale,
and reports the equivalent `q_hat` and `se_cr1`. Parametric AR(1), i.i.d., and
a positive user-supplied `psi` remain explicit sensitivity routes. The sign of
the cluster adjustment is never assumed.

The heterogeneity pilot is covariance-aware. The cluster-score covariance of
the group-time effects supplies the projected quadratic-form noise trace,
which is removed before the descriptive point radius is formed. Because that
quadratic pilot is nonregular at zero heterogeneity, it does not determine the
formal decision. Instead, cluster multipliers calibrate a one-sided confidence
radius for the projected group-time vector before its norm is taken. HC2 is the
primary bound and HC3 is reported as a leverage sensitivity check. This also
respects covariance across group-time comparisons that share controls.

```{r}
delta_g <- c(-0.6, 0, 0.6)                       # heterogeneous cohort effects
eff <- delta_g[ifelse(unit <= 16, 1, ifelse(unit <= 32, 2, 3))]
D <- as.numeric(!is.na(ft_stag) & time >= ft_stag)
y <- 0.1 * unit / N + D * eff + rnorm(N * T, sd = 0.5)
twfe_adequacy(y, unit, time, ft_stag)
```

The internal group-time estimator uses either not-yet-treated units (the
default, including never treated) or never-treated units only. Select the
published comparison group with `controls = "not_yet"` or
`controls = "never"`. This is a diagnostic implementation, not a replacement
for a full heterogeneity-robust DiD estimator.

## The applications, reproduced from bundled data

The package includes the exact four-column analysis extract for the
Callaway--Sant'Anna minimum-wage application, along with the castle-doctrine
and divorce panels. The minimum-wage data contain 15,988 county-years from
2,284 counties over seven years. The current diagnostics are:

- minimum wage: an equally weighted treated-cell ATT of about -5.2%, a 31.6%
  saturated-class point envelope, and a projected-norm lower endpoint above the
  threshold, so uniform certification is withheld;
- castle doctrine: a 5.1% saturated-class point envelope and an HC2 upper
  endpoint below the threshold, hence formal certification at `gamma = 0.05`;
  the slightly larger HC3 endpoint is a deliberately visible sensitivity;
- no-fault divorce: a roughly 70% saturated-class point envelope and a wide
  projected-norm interval crossing the threshold. The width is consistent with
  the strained fixed-`T` approximation (`G = 49`, `T = 27`).

The last result illustrates why the point envelope and its uncertainty should
both be reported. The package never relabels a worst-case envelope as realized
size.

```{r applications, eval=FALSE}
data(minimum_wage)
mw <- twfe_adequacy(minimum_wage$y, minimum_wage$uid,
                    minimum_wage$tid, minimum_wage$ft,
                    controls = "never", bootstrap = 999)
c(target_att = mw$statistic$att_target,
  envelope = mw$statistic$size_gt,
  K_lower = mw$statistic$K_lower_gt,
  directional = mw$statistic$size_directional,
  verdict = mw$verdict)

# Same protocol, opposite certification outcomes.
castle_report <- twfe_adequacy(castle$y, castle$uid, castle$tid, castle$ft)
divorce_report <- twfe_adequacy(divorce$y, divorce$uid, divorce$tid, divorce$ft)
c(castle = castle_report$verdict, divorce = divorce_report$verdict)
```

`data(package = "panelcert")` lists the bundled datasets. The full numerical
audit is also executable through the package tests and the sibling
`reproduction/reproduce_with_package.jl` harness.
