---
title: "Module C: TWFE-heterogeneity adequacy for staggered DiD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Module C: TWFE-heterogeneity adequacy for staggered DiD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed (see panelcert.Rmd).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(4)
```

## The design statistic, before any outcome exists

Under heterogeneous treatment effects, the TWFE coefficient is a weighted
average with weights `w` that need not be convex. Paper C shows the *size
distortion* of the TWFE t-test is governed by

`Gamma = sqrt(N1) * ||w - u||`,

the scaled distance of the weights from uniform — a **pure design statistic**,
computable from the adoption pattern alone. `Gamma = 0` exactly for clean
block designs; it grows with staggering and negative weights. Crucially,
`Gamma` is invariant to replicating the panel while the absolute weighting
bias shrinks — so a design can look safe to estimand-level diagnostics while
its t-test stays arbitrarily distorted.

```{r}
library(panelcert)
N <- 48; T <- 12
unit <- rep(1:N, each = T); time <- rep(1:T, times = N)

# a block design: one adoption date, never-treated reservoir
ft_block <- ifelse(unit <= 16, 5, NA)
twfe_design(unit, time, ft_block)$statistic$Gamma

# three cohorts, no never-treated reservoir (Paper C's simulation design)
ft_stag <- ifelse(unit <= 16, 3, ifelse(unit <= 32, 7, 11))
twfe_design(unit, time, ft_stag)
```

Run this before committing to a design: the report's breakdown value is the
largest heterogeneity-to-noise ratio `c/sigma` the design tolerates.

## The inference layer

With an outcome, `twfe_adequacy()` adds the realized non-centrality (from
cohort-level effect deviations under the TWFE weights) and the worst-case one
`(c/sigma) * Gamma`. Two honesty defaults are built in:

- the cohort-dispersion pilot is **shrinkage-corrected**:
  `var_shrunk = max(0, var(ATT_g) - mean(se_g^2))` — naive dispersion is
  estimation-noise-inflated and overstates the alarm;
- the cluster variance-inflation factor `psi` is **computed on the realized
  design** (parametric AR(1) route by default, with an estimator-driven
  cross-check), never assumed: persistent errors give `psi > 1` and shrink the
  distortion, exchangeable ones give `psi < 1` and worsen it.

```{r}
delta_g <- c(-0.6, 0, 0.6)                       # heterogeneous cohort effects
eff <- delta_g[ifelse(unit <= 16, 1, ifelse(unit <= 32, 2, 3))]
D <- as.numeric(!is.na(ft_stag) & time >= ft_stag)
y <- 0.1 * unit / N + D * eff + rnorm(N * T, sd = 0.5)
twfe_adequacy(y, unit, time, ft_stag)
```

The internal cohort pilot is a simplified not-yet-treated difference in means,
deliberately *not* a Callaway–Sant'Anna implementation; for a paper-grade
pilot, pass `cohort_effects` and `cohort_ses` from `did::att_gt()` /
`csdid` / `fixest::sunab` — the diagnostic then uses your estimates verbatim.
When a design is flagged, those same heterogeneity-robust estimators are the
remedy.

## The empirical poles (Paper C), reproduced from bundled data

Both audit panels ship with the package — `castle` and `divorce` — so the
two-pole result runs directly. Same estimator, same protocol, opposite
verdicts:

```{r}
# Castle doctrine: Gamma = 0.21, certified at size 5.0%, robust to clustering
twfe_adequacy(castle$y, castle$uid, castle$tid, castle$ft)$verdict

# No-fault divorce: Gamma = 0.86, flagged (60% iid -> 42% clustered)
d <- twfe_adequacy(divorce$y, divorce$uid, divorce$tid, divorce$ft)
c(Gamma = round(d$statistic$Gamma, 2), psi = round(d$statistic$psi_hat, 2),
  size = round(d$implied_size, 2), verdict = d$verdict)
```

The cluster inflation `psi = 1.62` is computed from the realized design, not
assumed; the full audit is reproduced by the package test suite. `data(package
= "panelcert")` lists the bundled datasets.
