---
title: "TWFE-heterogeneity adequacy for staggered DiD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TWFE-heterogeneity adequacy for staggered DiD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Evaluate chunks only when the package is actually installed (see panelcert.Rmd).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      eval = requireNamespace("panelcert", quietly = TRUE))
set.seed(4)
```

## The design statistic, before any outcome exists

Under heterogeneous treatment effects, the TWFE coefficient is a weighted
average with weights `w` that need not be convex. The accompanying article shows the *size
distortion* of the TWFE t-test is governed by

`Gamma = sqrt(N1) * ||w - u||`,

the scaled distance of the weights from uniform — a **pure design statistic**,
computable from the adoption pattern alone. `Gamma = 0` exactly for clean
block designs; it grows with staggering and negative weights. Crucially,
`Gamma` is invariant to replicating the panel while the absolute weighting
bias shrinks — so a design can look safe to estimand-level diagnostics while
its t-test stays arbitrarily distorted.

```{r}
library(panelcert)
N <- 48; T <- 12
unit <- rep(1:N, each = T); time <- rep(1:T, times = N)

# a block design: one adoption date, never-treated reservoir
ft_block <- ifelse(unit <= 16, 5, NA)
twfe_design(unit, time, ft_block)$statistic$Gamma

# three cohorts, no never-treated reservoir
ft_stag <- ifelse(unit <= 16, 3, ifelse(unit <= 32, 7, 11))
twfe_design(unit, time, ft_stag)
```

Run this before committing to a design: the report's breakdown value is the
largest heterogeneity-to-noise ratio `c/sigma` the design tolerates.

## The inference layer

With an outcome, `twfe_adequacy()` estimates the heterogeneity scale in the
cohort-plus-event-time class and maps it to two distinct quantities:

- the **worst-case size envelope** maximizes the mean shift over that calibrated
  class. It says what the design can certify uniformly; it is an upper bound,
  not a point prediction of the test's realized rejection probability;
- the **directional plug-in** inserts the estimated signed group-time profile.
  It is orientation-sensitive corroboration, not a uniform certificate.

The default scale is direct CR1. The function forms the unit cluster scores,
sets `psi_direct` to their CR1 variance relative to the homoskedastic scale,
and reports the equivalent `q_hat` and `se_cr1`. Parametric AR(1), i.i.d., and
a positive user-supplied `psi` remain explicit sensitivity routes. The sign of
the cluster adjustment is never assumed.

The heterogeneity pilot is covariance-aware. A fixed-design cluster bootstrap
estimates the covariance matrix of the group-time effects, and the projected
quadratic-form noise trace is removed before the class radius is formed. This
matters because group-time comparisons share controls; subtracting only the
average marginal variance generally removes the wrong amount of noise.

```{r}
delta_g <- c(-0.6, 0, 0.6)                       # heterogeneous cohort effects
eff <- delta_g[ifelse(unit <= 16, 1, ifelse(unit <= 32, 2, 3))]
D <- as.numeric(!is.na(ft_stag) & time >= ft_stag)
y <- 0.1 * unit / N + D * eff + rnorm(N * T, sd = 0.5)
twfe_adequacy(y, unit, time, ft_stag)
```

The internal group-time estimator uses either not-yet-treated units (the
default, including never treated) or never-treated units only. Select the
published comparison group with `controls = "not_yet"` or
`controls = "never"`. This is a diagnostic implementation, not a replacement
for a full heterogeneity-robust DiD estimator.

## The applications, reproduced from bundled data

The package includes the exact four-column analysis extract for the
Callaway--Sant'Anna minimum-wage application, along with the castle-doctrine
and divorce panels. The minimum-wage data contain 15,988 county-years from
2,284 counties over seven years. The current point diagnostics are:

- minimum wage: a 29.3% worst-case envelope and a 27.8% directional plug-in;
- castle doctrine: a 5.0% envelope, hence a certificate at the 10% ceiling;
- no-fault divorce: a 41.5% point envelope and a 37.1% directional plug-in,
  with a wide bootstrap interval and a fixed-`T` warning (`G = 49`, `T = 27`).

The last result illustrates why the point envelope and its uncertainty should
both be reported. The package never relabels a worst-case envelope as realized
size.

```{r applications, eval=FALSE}
data(minimum_wage)
mw <- twfe_adequacy(minimum_wage$y, minimum_wage$uid,
                    minimum_wage$tid, minimum_wage$ft,
                    controls = "never", bootstrap = 999)
c(envelope = mw$statistic$size_cmb,
  directional = mw$statistic$size_directional,
  verdict = mw$verdict)

# Same protocol, opposite certification outcomes.
castle_report <- twfe_adequacy(castle$y, castle$uid, castle$tid, castle$ft)
divorce_report <- twfe_adequacy(divorce$y, divorce$uid, divorce$tid, divorce$ft)
c(castle = castle_report$verdict, divorce = divorce_report$verdict)
```

`data(package = "panelcert")` lists the bundled datasets. The full numerical
audit is also executable through the package tests and the sibling
`reproduction/reproduce_with_package.jl` harness.
